---
title: "Coursera Statistical Inference Week4 Assignment"
author: "JS"
date: "March 13, 2017"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='figures/',warning=FALSE, message=FALSE,cache=TRUE)
```

# Question 1: 
Investigate the exponential distribution in R and compare it with the CLT. Exponential distribution can be simulated in R with rexp(n,lambda) 
where lambda is the rate parameter. The mean of exponential distribution is 1/lambda, and the SD is also 1/lambda. Set lambda = 0.2 for all the
simulations, we'll investigate the distribution of average of 40 exponentials, and do a thousand simulations. 

Illustrate the simulations and associated explanatory text the properties of the distribution of the mean of 40 exponentials. To do: 
* Show the sample mean and compare it to the theoretical mean of the distribution.
* Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
* Show that the distribution is approximately normal.

### Sample Mean vs Theoretical
```{r Q1: Sample Mean}
  
  #Simulate a thousand mean of exponential distribution, with parameter lambda:0.20
  #rexp generates random deviates of Exponential Distribution, the length of the result is determined by n parameter
  set.seed(100)

  lambda <- 0.2
  nbSimulations <- 1000
  nbPoints <- 40
    
      #Generate Matrix of 1000 Simulations of 40 points
      exp_simulated <- replicate(n=nbSimulations,expr=rexp(n=40,rate=lambda))
    
      #Calculate Sample mean of Generated Data
      exp_simulated_mean <- colMeans(exp_simulated)
      simulated_mean <- mean(exp_simulated)
      print(paste("The sample of mean of the 1000 Simulations is ",simulated_mean, " compared to the theoretical mean of ", 1/lambda ))

```

### Sample Variance vs Theoretical
```{r Q2: Sample Variance}
  
  #Simulate a thousand mean of exponential distribution, with parameter lambda:0.20
  #rexp generates random deviates of Exponential Distribution, the length of the result is determined by n parameter

    
      #Calculate Sample Stdev of Generated Data
      simulated_Stdev <- sd(exp_simulated_mean)*sqrt(40)
      print(paste("The sample SD of the 1000 Simulations is ",simulated_Stdev, " compared to the theoretical SD of ", 1/lambda ))

```

### Normal Distribution comparison 
  We can observe on the below QQ plot that the sample Quantiles are very close to the theoretical, same as the Density function. 

```{r Q3: Distribution}
 #Plot Theoretical Quantiles vs Sample Quantiles
  qqnorm(exp_simulated_mean);qqline(exp_simulated_mean,col=2)
 #Distribution
    #Theoretical 
      x_theo <- seq(min(exp_simulated_mean),max(exp_simulated_mean),length=nbPoints)
      y_theo <- dnorm(x_theo,mean=1/lambda,sd=(1/lambda/sqrt(nbPoints)))

#Histogram 
 hist(exp_simulated_mean,breaks=nbPoints,prob=T,col="orange",xlab = "means",ylab="density")
 lines(x_theo,y_theo)
                  
```

#Question 2: Basic Inferential Data Analysis

Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. To do: 

*Load the ToothGrowth data and perform some basic exploratory data analyses
*Provide a basic summary of the data.
*Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
*State your conclusions and the assumptions needed for your conclusions.


```{r Data Analysis}
  #Load Data set
    data("ToothGrowth")

  #Get Information about the data Set
    ?ToothGrowth

  #Here's what we can learn from the help page: 
    #The response is the length of odontoblasts (cells responsible for tooth growth) 
    #in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods,
    #(orange juice or ascorbic acid (a form of vitamin C and coded as VC).

  #Show Type of DataSet
    str(ToothGrowth)
  #Show Sample
    head(ToothGrowth)
  #Summary 
    summary(ToothGrowth)
    
  #View the histogram of length by dose, on each delivery methods  
    ggplot(ToothGrowth,aes(x=len,fill=supp))+facet_wrap(~dose)+geom_histogram()
    
  #Performing T-test on each dose
  
```
On the three histograms, we can observe that for 0.5 and 1mg/day doses it seems like Orange Juice treatment get higher length, whereas for 2mg/day Vitamin C gets higher.

Now let's perform a T-Test analysis on supp: 

```{r T-test performing}
  d_05 <- ToothGrowth[ToothGrowth$dose=="0.5",]
  t.test(data=d_05,len ~ supp,paired=FALSE,var.equal=FALSE)
  
  d_1 <- ToothGrowth[ToothGrowth$dose=="1",]
  t.test(data=d_1,len ~ supp,paired=FALSE,var.equal=FALSE)
  
  d_2 <- ToothGrowth[ToothGrowth$dose=="2",]
  t.test(data=d_2,len ~ supp,paired=FALSE,var.equal=FALSE)

```

* 0.5mg/day dose
95% confidence interval does not contain zero, we can reject null hypothesis. The tooth growth for OJ is statistically significantly higher than VC treatment. 

* 1mg/day dose 
95% confidence interval does not contain zero, we can reject null hypothesis. The tooth growth for OJ is statistically significantly higher than VC treatment. 

* 2mg/day dose 
95% confidence interval contains zero, we can not reject the null hypothesis. Then, at the highest dose, there is no difference in tooth growth between two treatments. 

Then, let's see if on each treatment the fact of increasing the dosage increase the tooth growth: 
 OJ : 

```{r T-test performing2}
  OJ_Min <- ToothGrowth[ToothGrowth$dose=="0.5" & ToothGrowth$supp=="OJ",]
  OJ_Med <- ToothGrowth[ToothGrowth$dose=="1" & ToothGrowth$supp=="OJ",]
  OJ_Max <- ToothGrowth[ToothGrowth$dose=="2" & ToothGrowth$supp=="OJ",]
  
  #Compare Min with Med
  t.test(OJ_Min$len,OJ_Med$len,paired=FALSE,var.equal = FALSE)
  
  #Compare Max with Med 
  t.test(OJ_Max$len,OJ_Med$len,paired=FALSE,var.equal = FALSE)
```
* When comparing OJ treatment at 0.5 and 1 mg dosages, 95% confidence does not contain zero, then increasing the dosage of OJ increase tooth growth. 
* When comparing OJ treatment at 1mg and 2 mg dosages, 95% confidence does not contain zero, then increasing the dosage of OJ increase tooth growth. 

Let's do the same analysis on VC. 

```{r T-test performing2}
  VC_Min <- ToothGrowth[ToothGrowth$dose=="0.5" & ToothGrowth$supp=="VC",]
  VC_Med <- ToothGrowth[ToothGrowth$dose=="1" & ToothGrowth$supp=="VC",]
  VC_Max <- ToothGrowth[ToothGrowth$dose=="2" & ToothGrowth$supp=="VC",]
  
  #Compare Min with Med
  t.test(VC_Min$len,VC_Med$len,paired=FALSE,var.equal = FALSE)
  
  #Compare Max with Med 
  t.test(VC_Max$len,VC_Med$len,paired=FALSE,var.equal = FALSE)
```

* We can observe the same conclusion on VC treatment, with an increasing dosage of VC that increase tooth growth. 

## Conclusion

 * Both with VC and OJ treatment, the fact to increase the dosage significantly increase tooth growth.
 * At 0.5 and 1mg dosages, OJ treatment have significantly higher tooth growth than VC
 * At 2mg dosage, no significant different tooth growth 
 


